---
title: "ProjectProposal"
author: "Michael Thane"
date: "11 5 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Research Questions
- How does the popularity of programming languages change over time?
  - How can we measure popularity of a programming language?
  - Number of upvotes, comments
  - Type of language

- Is a certain post refering to a professional question?
  - How can we classify those questions?
  - Sentiment Analysis? 
  - Question in the title

- Is a certain post refering to an opinion?
  - Opinion Mining, Sentiment Analysis
- What programming languages are used for certain problems or programming paradigm?
  - Topic modeling
 
- Which programming languages are important for certain jobs?

## Reddit Scraping
### Data set example


```{r reddit, echo=FALSE,warning=FALSE,message=FALSE}
library(dplyr)
library(jsonlite)
       fromJSON(url(paste("https://www.reddit.com/r/",
                          "coding",
                          "/new/.json?limit=",
                          10,sep="")),flatten = T)$data$children%>%
                      select("data.subreddit",
                             "data.title",
                             "data.author",
                             "data.selftext",
                             "data.name",
                             "data.id",
                             "data.domain",
                             "data.url",
                             "data.created",
                             "data.created_utc",
                             "data.upvote_ratio",
                             "data.ups",
                             "data.score",
                             "data.num_comments"
                             )%>%
         knitr::kable()
```

## Design overview

### How can we measure popularity of a programming language?
- upvotes and number of columns

### How can we tell what language a post is referring to?
- using a ground truth to predict what programming language a post is related to
- Using topic modeling using LDA - Latent Dirichlet Allocation
- Using decision trees or ensemble methods

### Visualization
- Showing the trend with a line plot and confidence interval
- Using 2D embedding of the posts and using a scatterplot to show similarity 
- Comparing stackoverflow and Reddit in the plots

### Is a certain post referring to a professional question or an opinion?
- Using sentiment analysis and clustering 


## Time plan including distribution of responsibilities and workload among team members written as weekly deadlines

Data processing - Reddit | Michael & Sharanya
Data processing - Stackoverflow
Visualization and EDA | Michael & Indranil & Sharanya
Topic modeling | Sharanya & Indrani
Classification | 

Shiny application - Everyone