---
title: "ProjectProposal"
author: "Michael Thane"
date: "11 5 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Bullet points

- this is a collaborative document
- commit push and pull often, also when there is only small changes
- add your ideas in the following paragraph "ideas"

## Ideas
- Web scraping on Reddit
    - Alternative visualization of Reddit posts (2D embedding)
    - Sentiment Analysis
    - Explainable predictive Model
    - Predictive Model + visualization
- Rise and Fall of Programming Languages based on Stackoverflow data

## Research Questions
- How does the popularity of programming languages change over time?
  - How can we measure popularity of a programming language?
  - Number of upvotes, comments
  - Type of language

- Is a certain post refering to a professional question?
  - How can we classify those questions?
  - Sentiment Analysis? 
  - Question in the title

- Is a certain post refering to an opinion?
  - Opinion Mining, Sentiment Analysis
- What programming languages are used for certain problems or programming paradigm?
  - Topic modeling
 
- Which programming languages are important for certain jobs?

## Reddit Scraping
### Data set example


```{r reddit, echo=FALSE,warning=FALSE,message=FALSE}
library(dplyr)
library(jsonlite)
       fromJSON(url(paste("https://www.reddit.com/r/",
                          "coding",
                          "/new/.json?limit=",
                          10,sep="")),flatten = T)$data$children%>%
                      select("data.subreddit",
                             "data.title",
                             "data.author",
                             "data.selftext",
                             "data.name",
                             "data.id",
                             "data.domain",
                             "data.url",
                             "data.created",
                             "data.created_utc",
                             "data.upvote_ratio",
                             "data.ups",
                             "data.score",
                             "data.num_comments"
                             )%>%
         knitr::kable()
```

## Programming Challenges


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
